---
title: "Demographic and Economic Benchmarks"
author: "Kassie Breest, Max Williams"
date: "2024-08-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Read in Packages

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(purrr)
library(janitor)
library(flextable)
library(officer)
library(lubridate)
library(openxlsx)
library(rio)
library(dplyr)
library(haven)
library(tidycensus)
```

#TidyCensus Set Up
Need town level data for ACS 5YR 2018-2022 in MA, Essex County, and the 9 Municipalities in the study

ACS variables to pull: 
- population: TOTAL POPULATION = B01003 
- median household income: HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2021 INFLATION-ADJUSTED DOLLARS) = B19013 
- educational distribution: EDUCATIONAL ATTAINMENT = S1501 
- racial distribution: RACE = B02001 
- ethnic distribution?: HISPANIC OR LATINO ORIGIN = B03003 
- age distribution: AGE AND SEX = S0101

##Setting up variables and geographies

```{r}
towns <- "County Subdivision"
counties <- "County"
years <- c(2022:2012)
outputs <- "tidy"
states <- 25
tables <- c("B01003", "B19013", "S1501", "B02001", "B03003", "S0101")
```


##Study Towns
```{r}
study_comm <- read_excel("StudyCommunities.xlsx")

study_comm <- study_comm %>%
  rename(GEOID = GeoID)

```

## Investigate the data to plan for cleaning

```{r}
varnames_subject <- load_variables(2018, "acs5/subject") #looking for 2017-2021 variables
View(varnames_subject)

varnames_detailed <- load_variables(2018, "acs5")
View(varnames_detailed)
```

#ACS
##Total Population
```{r}
#TOTAL POPULATION
studytowns_pop <- map_df(years, function(y) {
  get_acs(
    geography = towns, 
    state = states,
    table = "B01003",
    year = y,
    geometry = FALSE,
    output = outputs,
    cache_table = TRUE
  ) %>%
    mutate(year = y)
}) 

studytowns_pop <- studytowns_pop %>%
  mutate(GEOID = as.numeric(GEOID))

studytowns_pop <- left_join(study_comm, studytowns_pop, by = "GEOID")

#Population
fortable_pop <- studytowns_pop %>%
  group_by(year) %>%
  rename(pop = estimate) %>% #renamed this variable to avoid confusion
  select(year, place, state, GEOID, ES202_match, pop) %>% #select only the necessary variables for the table
  distinct(place, .keep_all = TRUE) 

```

##Median Household Income
```{r}
#HOUSEHOLD INCOME
studytowns_HHinc <- map_df(years, function(y) {
  get_acs(
    geography = towns, 
    state = states,
    table = "B19013",
    year = y,
    geometry = FALSE,
    output = outputs,
    cache_table = TRUE
  ) %>%
    mutate(year = y)
}) 

studytowns_HHinc <- studytowns_HHinc %>%
  mutate(GEOID = as.numeric(GEOID))

studytowns_HHinc <- left_join(study_comm, studytowns_HHinc, by = "GEOID")

#Median Household Income 
fortable_medianHHinc <- studytowns_HHinc %>%
  group_by(year) %>%
  rename(medianHHinc = estimate) %>% #renamed this variable to avoid confusion
  select(year, place, state, GEOID, ES202_match, medianHHinc) %>% #select only the necessary variables for the table
  distinct(place, .keep_all = TRUE) 
```

##Educational Distribution
```{r}
#EDUCATIONAL ATTAINMENT
studytowns_edu <- map_df(years, function(y) {
  get_acs(                                                                                                
    geography = towns, 
    state = states,
    table = "S1501",
    year = y,
    geometry = FALSE,
    output = outputs,
    cache_table = TRUE
  ) %>%
    mutate(year = y)
}) 

studytowns_edu <- studytowns_edu %>%
  mutate(GEOID = as.numeric(GEOID)) #fix GEOID for merge

edu_vars <- read_excel("EduAttain_S1501_SelectedVars.xlsx") #read in file of selected variables

studytowns_edu_joined <- studytowns_edu %>%
  inner_join(edu_vars, by = "variable") #inner join to select only the values of the variable "variable" that we want (drops obs)

studytowns_edu_joined <- left_join(study_comm, studytowns_edu_joined, by = "GEOID") #left join to select only the values of the variable GEOID that are in our study

fortable_edu <- studytowns_edu_joined %>%
  pivot_wider(
    id_cols = c(GEOID, NAME, year, place, muni, state, ES202_match, BLS_match), 
    names_from = variable_name, 
    values_from = estimate
  ) #pivot wider so that there is one variable for each value of the variable "variable", named using names specified in edu_vars

```

##Race
```{r}
#RACE
studytowns_race <- map_df(years, function(y) {
  get_acs(
    geography = towns, 
    state = states,
    table = "B02001",
    year = y,
    geometry = FALSE,
    output = outputs,
    cache_table = TRUE
  ) %>%
    mutate(year = y)
}) 

studytowns_race <- studytowns_race %>%
  mutate(GEOID = as.numeric(GEOID))

race_vars <- read_excel("Race_B02001_SelectedVars.xlsx") #read in file of selected variables

studytowns_race_joined <- studytowns_race %>%
  inner_join(race_vars, by = "variable") #inner join to select only the values of the variable "variable" that we want (drops obs)

studytowns_race_joined <- left_join(study_comm, studytowns_race_joined, by = "GEOID") #left join to select only the values of the variable GEOID that are in our study

fortable_race <- studytowns_race_joined %>%
  pivot_wider(
    id_cols = c(GEOID, NAME, year, place, muni, state, ES202_match, BLS_match), 
    names_from = variable_name, 
    values_from = estimate
  ) #pivot wider so that there is one variable for each value of the variable "variable", named using names specified in race_vars

```

##Ethnicity
```{r}
#HISPANIC OR LATINO ORIGIN
studytowns_hispanic <- map_df(years, function(y) {
  get_acs(
    geography = towns, 
    state = states,
    table = "B03003",
    year = y,
    geometry = FALSE,
    output = outputs,
    cache_table = TRUE
  ) %>%
    mutate(year = y)
}) 

studytowns_hispanic <- studytowns_hispanic %>%
  mutate(GEOID = as.numeric(GEOID))

hispanic_vars <- read_excel("Hispanic_B03003_SelectedVars.xlsx") #read in file of selected variables

studytowns_hispanic_joined <- studytowns_hispanic %>%
  inner_join(hispanic_vars, by = "variable") #inner join to select only the values of the variable "variable" that we want (drops obs)

studytowns_hispanic_joined <- left_join(study_comm, studytowns_hispanic_joined, by = "GEOID") #left join to select only the values of the variable GEOID that are in our study

fortable_hispanic <- studytowns_hispanic_joined %>%
  pivot_wider(
    id_cols = c(GEOID, NAME, year, place, muni, state, ES202_match, BLS_match), 
    names_from = variable_name, 
    values_from = estimate
  ) #pivot wider so that there is one variable for each value of the variable "variable", named using names specified in race_vars

```

##Age 
```{r}
#AGE AND SEX
studytowns_agesex <- map_df(years, function(y) {
  get_acs(
    geography = towns, 
    state = states,
    table = "S0101",
    year = y,
    geometry = FALSE,
    output = outputs,
    cache_table = TRUE
  ) %>%
    mutate(year = y)
}) 

studytowns_agesex <- studytowns_agesex %>%
  mutate(GEOID = as.numeric(GEOID))

age_vars <- read_excel("Age_S0101_SelectedVars.xlsx") #read in file of selected variables

studytowns_age_joined <- studytowns_agesex %>%
  inner_join(age_vars, by = "variable") #inner join to select only the values of the variable "variable" that we want (drops obs)

studytowns_age_joined <- left_join(study_comm, studytowns_age_joined, by = "GEOID") #left join to select only the values of the variable GEOID that are in our study

fortable_age <- studytowns_age_joined %>%
  pivot_wider(
    id_cols = c(GEOID, NAME, year, place, muni, state, ES202_match, BLS_match), 
    names_from = variable_name, 
    values_from = estimate
  ) #pivot wider so that there is one variable for each value of the variable "variable", named using names specified in race_vars

fortable_age_2017_2022 <- fortable_age %>%
  filter(year > 2016) %>%
  mutate(
    pct_under_5 = (total_under_5 / total) * 100,
    pct_5_to_9 = (total_5_to_9 / total) * 100,
    pct_10_to_14 = (total_10_to_14 / total) * 100,
    pct_15_to_19 = (total_15_to_19 / total) * 100,
    pct_20_to_24 = (total_20_to_24 / total) * 100,
    pct_25_to_29 = (total_25_to_29 / total) * 100,
    pct_30_to_34 = (total_30_to_34 / total) * 100,
    pct_35_to_39 = (total_35_to_39 / total) * 100,
    pct_40_to_44 = (total_40_to_44 / total) * 100,
    pct_45_to_49 = (total_45_to_49 / total) * 100,
    pct_50_to_54 = (total_50_to_54 / total) * 100,
    pct_55_to_59 = (total_55_to_59 / total) * 100,
    pct_60_to_64 = (total_60_to_64 / total) * 100,
    pct_65_to_69 = (total_65_to_69 / total) * 100,
    pct_70_to_74 = (total_70_to_74 / total) * 100,
    pct_75_to_79 = (total_75_to_79 / total) * 100,
    pct_80_to_84 = (total_80_to_84 / total) * 100,
    pct_85_over = (total_85_over / total) * 100,
    pct_selected_5_to_14 = (selected_5_to_14 / total) * 100,
    pct_selected_15_to_17 = (selected_15_to_17 / total) * 100,
    pct_selected_under_18 = (selected_under_18 / total) * 100,
    pct_selected_18_to_24 = (selected_18_to_24 / total) * 100,
    pct_selected_15_to_44 = (selected_15_to_44 / total) * 100
  )

fortable_age_2012_2016 <- fortable_age %>%
  filter(year < 2016) %>%
  rename_with(
    .fn = ~ gsub("^total_", "pct_", .),  # Rename 'total_' to 'pct_'
    .cols = starts_with("total_")
  ) %>%
  rename_with(
    .fn = ~ paste0("pct_", .),  # Add 'pct_' as prefix to 'selected_' columns
    .cols = starts_with("selected_")
  )%>%
  mutate(
    total_under_5 = (pct_under_5 / 100) * total,
    total_5_to_9 = (pct_5_to_9 / 100) * total,
    total_10_to_14 = (pct_10_to_14 / 100) * total,
    total_15_to_19 = (pct_15_to_19 / 100) * total,
    total_20_to_24 = (pct_20_to_24 / 100) * total,
    total_25_to_29 = (pct_25_to_29 / 100) * total,
    total_30_to_34 = (pct_30_to_34 / 100) * total,
    total_35_to_39 = (pct_35_to_39 / 100) * total,
    total_40_to_44 = (pct_40_to_44 / 100) * total,
    total_45_to_49 = (pct_45_to_49 / 100) * total,
    total_50_to_54 = (pct_50_to_54 / 100) * total,
    total_55_to_59 = (pct_55_to_59 / 100) * total,
    total_60_to_64 = (pct_60_to_64 / 100) * total,
    total_65_to_69 = (pct_65_to_69 / 100) * total,
    total_70_to_74 = (pct_70_to_74 / 100) * total,
    total_75_to_79 = (pct_75_to_79 / 100) * total,
    total_80_to_84 = (pct_80_to_84 / 100) * total,
    total_85_over = (pct_85_over / 100) * total,
    selected_5_to_14 = (pct_selected_5_to_14 / 100) * total,
    selected_15_to_17 = (pct_selected_15_to_17 / 100) * total,
    selected_under_18 = (pct_selected_under_18 / 100) * total,
    selected_18_to_24 = (pct_selected_18_to_24 / 100) * total,
    selected_15_to_44 = (pct_selected_15_to_44 / 100) * total
  )

fortable_age_2012_2016 <- fortable_age_2012_2016[names(fortable_age_2017_2022)]
fortable_age_2017_2022 <- fortable_age_2017_2022[names(fortable_age_2012_2016)]

fortable_age <- rbind(fortable_age_2012_2016, fortable_age_2017_2022)
```


#Figures

