---
title: "Community Comparisons FY24"
author: "Melissa, Barbara Talagan"
date: "2023-08-21"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

#Load packages

```{r message=FALSE, warning=FALSE}
# For Census Data
library(tidycensus)
library(tidyverse)
library(foreign)
library(openxlsx)
library(writexl)
library(flextable)
library(officer)
library(readr)
library(readxl)
library(purrr)
library(janitor)
library(dplyr)
library(ggplot2)
library(scales)
library(tools)
library(stringr)
library(extrafont)


```
#ACS
##Pull data 
Need town level data for ACS 5YR 2017-2021, 9 states (ME, NH,
VT, MA, CT, RI, NY, PA, NJ)

ACS variables to pull: 
- population: TOTAL POPULATION = B01003 
- median household income: HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2021 INFLATION-ADJUSTED DOLLARS) = B19013 
- percent of pop over 25 with only bachelor's degree: EDUCATIONAL ATTAINMENT = S1501 
- percent of pop, black: RACE = B02001 
- percent of pop, hispanic: HISPANIC OR LATINO ORIGIN = B03003 
- percent of pop below poverty: POVERTY STATUS IN THE PAST 12 MONTHS = S1701 
- working age pop (percent of pop between 16-64): AGE AND SEX = S0101

##Setting up variables and geographies

```{r}
towns <- "County Subdivision"
years <- 2021
outputs <- "tidy"
#states <- c(09, 25, 23, 33, 34, 36, 44, 50, 42)
states <- c(09,25,33,36)
tables <- c("B01003", "B19013", "S1501", "B02001", "B03003", "S1701", "S0101")
```

```{r}
study_comm <- read_excel("StudyCommunities.xlsx")

study_comm <- study_comm %>%
  rename(GEOID = GeoID)

```


##Create one df for each table pulled

```{r}
#TOTAL POPULATION
studytowns_pop <- get_acs(
  geography = towns,
  state = states,
  table = "B01003",
  year = years,
  geometry = FALSE,
  output = outputs,
  cache_table = TRUE
)

studytowns_pop <- left_join(study_comm, studytowns_pop, by = "GEOID")


#HOUSEHOLD INCOME
studytowns_HHinc <- get_acs(
  geography = towns,
  state = states,
  table = "B19013",
  year = years,
  geometry = FALSE,
  output = outputs,
  cache_table = TRUE
)

studytowns_HHinc <- left_join(study_comm, studytowns_HHinc, by = "GEOID")

#EDUCATIONAL ATTAINMENT
studytowns_edu <- get_acs(
  geography = towns,
  state = states,
  table = "S1501",
  year = years,
  geometry = FALSE,
  output = outputs,
  cache_table = TRUE
)

studytowns_edu <- left_join(study_comm, studytowns_edu, by = "GEOID")

#RACE
studytowns_race <- get_acs(
  geography = towns,
  state = states,
  table = "B02001",
  year = years,
  geometry = FALSE,
  output = outputs,
  cache_table = TRUE
)

studytowns_race <- left_join(study_comm, studytowns_race, by = "GEOID")

#HISPANIC OR LATINO ORIGIN
studytowns_hispanic <- get_acs(
  geography = towns,
  state = states,
  table = "B03003",
  year = years,
  geometry = FALSE,
  output = outputs,
  cache_table = TRUE
)

studytowns_hispanic <- left_join(study_comm, studytowns_hispanic, by = "GEOID")

#POVERTY STATUS
studytowns_povertystat <- get_acs(
  geography = towns,
  state = states,
  table = "S1701",
  year = years,
  geometry = FALSE,
  output = outputs,
  cache_table = TRUE
)

studytowns_povertystat <- left_join(study_comm, studytowns_povertystat, by = "GEOID")

#AGE AND SEX
studytowns_agesex <- get_acs(
  geography = towns,
  state = states,
  table = "S0101",
  year = years,
  geometry = FALSE,
  output = outputs,
  cache_table = TRUE
)

studytowns_agesex <- left_join(study_comm, studytowns_agesex, by = "GEOID")
```


## Investigate the data to plan for cleaning

```{r}
varnames_subject <- load_variables(2017, "acs5/subject") #looking for 2017-2021 variables
View(varnames_subject)

varnames_detailed <- load_variables(2017, "acs5")
View(varnames_detailed)
```

## Clean data
###1. Population

```{r}
fortable_pop <- studytowns_pop %>%
  select(place, comm_type, state, GEOID, estimate) %>%
  rename(pop_1 = estimate)%>%
  distinct(place, .keep_all = TRUE)

```



###2. Percent of Pop. over 25 with only a BS
```{r}
#Edu. Attain College as % of pop 25 and over or Percent BS only

#POP 25 AND OVER
studytowns_pop25 <- studytowns_agesex %>%
  filter(variable %in% c(
    "S0101_C01_008", "S0101_C01_009", "S0101_C01_010", "S0101_C01_011",
    "S0101_C01_012", "S0101_C01_013", "S0101_C01_014", "S0101_C01_015",
    "S0101_C01_016", "S0101_C01_017", "S0101_C01_018", "S0101_C01_019"
  )) %>% #filter for all age groups over 25
  group_by(GEOID, place, comm_type, state) %>%
  summarize(pop25 = sum(estimate)) #sum each age group count by community to get the total pop over 25
  

#POP WITH A BS
#Look at the varnames dataframe, filter for "S1501" and scroll through until you locate the variable for "Estimate of Total Population 25 years and over with a Bachelor's degree". The variable you want is "S1501_C01_012", so filter the dataframe for that variable and now you'll have what we need

fortable_pctbsonly <- studytowns_edu %>%
  filter(variable == "S1501_C01_012") %>%  #filter for only the estimates of pop 25 yrs and over with a BS
  rename(popwithBS = estimate) #renamed this variable to avoid confusion

fortable_pctbsonly<- left_join(fortable_pctbsonly, studytowns_pop25, by = "GEOID") %>% #add in the total pop variable
  select(place.x, comm_type.x, state.x, GEOID, pop25, popwithBS) %>%
  rename(state = state.x,
         place = place.x,
         comm_type = comm_type.x) %>%
  mutate(pctbs_2 = (popwithBS/pop25)*100)%>%
  distinct(place, .keep_all = TRUE) #calculate % of pop that have a BS

```

###3. Median Household Income
```{r}  
#Median Household Income 
fortable_medianHHinc <- studytowns_HHinc %>%
  filter(variable == "B19013_001") %>%  #filter for overall median household income
  rename(medianHHinc_3 = estimate) %>% #renamed this variable to avoid confusion
  select(place, comm_type, state, GEOID, medianHHinc_3) %>% #select only the necessary variables for the table
  distinct(place, .keep_all = TRUE) 

```


###4. Percent of Pop. that is Black
```{r}
#Percent of population that is black
studytowns_pctblack <- studytowns_race %>%
  filter(variable == "B02001_003") %>% #filter for only black population
  rename(popblack = estimate) #renamed this variable to avoid confusion

fortable_pctblack <- left_join(studytowns_pctblack, fortable_pop, by = "GEOID") %>% #add in the total pop variable
  select(place.x, comm_type.x, state.x, GEOID, pop_1, popblack) %>%
  rename(state = state.x,
         place = place.x,
         comm_type = comm_type.x) %>%
  mutate(pctblack_4 = (popblack/pop_1)*100)%>%#calculate % of pop that is black
  distinct(place, .keep_all = TRUE) 

```

###5. Percent of Pop. that is Hispanic
```{r}
#Percent of population that is hispanic/latino
studytowns_pcthispanic <- studytowns_hispanic %>%
  filter(variable == "B03003_003") %>% #filter for only hispanic/latino population
  rename(pophispanic = estimate) #renamed this variable to avoid confusion

fortable_pcthispanic <- left_join(studytowns_pcthispanic, fortable_pop, by = "GEOID")%>% #add in the total pop variable
  select(place.x, comm_type.x, state.x, GEOID, pop_1, pophispanic) %>%
  rename(state = state.x,
         place = place.x,
         comm_type = comm_type.x) %>%
    mutate(pcthispanic_5 = (pophispanic/pop_1)*100) %>% #calculate % of pop that is hispanic/latino
  distinct(place, .keep_all = TRUE) 

```

###6. Percent of Pop. below Poverty
```{r}
#Percent of population in poverty
studytowns_pctpoverty <- studytowns_povertystat %>%
  filter(variable == "S1701_C02_001") %>% #filter for total population in poverty
  rename(poppoverty = estimate) #renamed this variable to avoid confusion

fortable_pctpoverty<- left_join(studytowns_pctpoverty, fortable_pop, by = "GEOID")  %>% #add in the total pop variable
    select(place.x, comm_type.x, state.x, GEOID, pop_1, poppoverty) %>%
    rename(state = state.x,
         place = place.x,
         comm_type = comm_type.x) %>%
  mutate(pctpoverty_6 = (poppoverty/pop_1)*100) %>% #calculate % of pop that is in poverty
  distinct(place, .keep_all = TRUE) 

```

### 7. Working Age Pop.
```{r}
#Percent of working age population
#Population over 16
studytowns_popover16 <- studytowns_agesex %>%
  filter(variable == "S0101_C01_025") %>% #filter for pop over 16
  rename(popover16 = estimate) #renamed variable to avoid confusion

##################################################################################
#Population over 65
#studytowns_popover65 <- studytowns_agesex %>%
  #filter(variable == "S0101_C01_030") %>% #filter for just population over 65 (labor force)
  #rename(popover65 = estimate) #renamed variable to avoid confusion

#Combining df then calculating pop btwn 16-64 and percentgae of working age pop
#studytowns_workingpop <- left_join(studytowns_popover16, studytowns_popover65, by = "GEOID")%>%
  #select(place.x, comm_type.x, state.x, BLS_match.x, GEOID, popover16, popover65) %>%
  #rename(state = state.x,
         #place = place.x,
         #comm_type = comm_type.x,
        # BLS_match = BLS_match.x) %>%
 # mutate(workingpop = popover16 - popover65) %>% #subtracting pop over 65 from pop over 16 to calculate pop between 16-64
  #distinct(place, .keep_all = TRUE) 
##################################################################################

#METHOD CHANGE 04.22.24: 16+ population as opposed to 16-64
studytowns_workingpop <- studytowns_popover16 %>%
  select(place, comm_type, state, BLS_match, GEOID, popover16) %>%
  mutate(workingpop = popover16) %>%
  select(-popover16)%>%
  distinct(place, .keep_all = TRUE) 

```

#BLS
##Import Data
```{r}
#MA Towns
#import
BLS_MA_path <-"BLS_LAUS_2013-2023/MA Towns.xlsx"
BLS_MA <- read_excel(BLS_MA_path)

#clean
BLS_MA <- left_join(BLS_MA, study_comm, by = c("place" = "BLS_match"))%>%
  filter(year >= 2017 & year <= 2021) %>%
  select(place.y, year, period, comm_type, muni, state, GEOID, NewGeoID, ES202_match, `labor force`, employment, unemployment, `unemployment rate`)%>%
  rename(place = place.y)


#Study Comm
#import
BLS_path <-"J:/EPPR/SEIGMA/Secondary Data/Census Data/Kassie/censusdata/BLS_LAUS_2013-2023/study comm/"
BLS_workbooks <- list.files(BLS_path)
BLSworkbook_names <-stringr::str_remove(BLS_workbooks, ".xlsx")

process1 <- lapply(paste(BLS_path, BLS_workbooks, sep = ""), read_excel)
names(process1) <- BLSworkbook_names

BLS <- bind_rows(process1) 

#clean
BLS <- BLS %>%
  mutate(Place = rep(BLSworkbook_names, each = nrow(BLS) / length(BLSworkbook_names)))%>%
  filter(Year >= 2017 & Year <= 2021) %>%
  rename(place = Place,
         year = Year,
         period = Period)

BLS <- left_join(BLS, study_comm, by = "place") %>%
  filter(state != "MA")

BLS_all <- full_join(BLS, BLS_MA, by = "place") %>%
    mutate(
      year = coalesce(year.y, year.x),
      period = coalesce(period.y, period.x),
      `labor force` = coalesce(`labor force.y`, `labor force.x`),
      employment = coalesce(employment.y, employment.x),
      unemployment = coalesce(unemployment.y, unemployment.x),
      `unemployment rate` = coalesce(`unemployment rate.y`, `unemployment rate.x`),
      comm_type = coalesce(comm_type.y, comm_type.x),
      muni = coalesce(muni.y, muni.x),
      state = coalesce(state.y, state.x),
      GEOID = coalesce(GEOID.y, GEOID.x),
      NewGeoID = coalesce(NewGeoID.y, NewGeoID.x),
      ES202_match = coalesce(ES202_match.y, ES202_match.x),
      # Add more variables as needed
      .keep = "unused"
    )
    

```


##Clean Data


###8. Labor Force Count
```{r}
#average labor force by town across 5 yr period
#studytowns_LF <- BLS_all %>%
  #group_by(place, year) %>%
  #summarize(lf_total = mean(`labor force`, na.rm = TRUE)) %>%
  #ungroup()%>%
  #select(place, lf_total) %>%
  #distinct(place, .keep_all = TRUE)

########################################################################################
#METHOD CHANGE 04.22.24: average by year first, then by place

#average labor force for each year, for each place (months -> years)
studytowns_LF <- BLS_all %>%
  group_by(place, year) %>%
  summarize(lf_total = mean(`labor force`, na.rm = TRUE)) 

#average labor force for each place (years -> single point average)
studytowns_LF <- studytowns_LF %>%
  group_by(place) %>%
  summarize(lf_total = mean(lf_total, na.rm = TRUE))%>%
  ungroup()%>%
  select(place, lf_total) %>%
  distinct(place, .keep_all = TRUE)


```


###8.2. Labor Force Participation Rate
7 and 8
```{r}
#LFPR = total labor force/working age population

fortable_lfpr <- left_join(studytowns_workingpop, studytowns_LF, by = "place") %>%
  mutate(lfpr_8.2 = (lf_total/workingpop)*100) %>%
  select(place, comm_type, state, GEOID, lf_total, workingpop, lfpr_8.2) %>%
  distinct(place, .keep_all = TRUE)

#fortable_lfpr <- left_join(fortable_lfpr, MAtowns_LF, by = "place") %>%
  #mutate(lf_total = ifelse(is.na(lf_total.y), lf_total.x, lf_total.y)) %>%


```


###9. Unemployment Rate
```{r}
#average unemp. rate by town across 5 year period

#fortable_unemp <- BLS_all %>% 
 # group_by(place, comm_type, state, GEOID) %>%
 # summarize(avgunemp_9 = mean(`unemployment rate`, na.rm = TRUE),
  #          .groups = "drop")%>%
 # distinct(place, .keep_all = TRUE)

########################################################################################
#METHOD CHANGE 04.22.24: create rate and average by year first, then by place

#average unemployment for each year, for each place (months -> years)
studytowns_UI <- BLS_all %>%
  group_by(place, year) %>%
  summarize(ui_total = mean(unemployment, na.rm = TRUE)) 

#average labor force for each place (years -> single point average)
studytowns_UI <- studytowns_UI %>%
  group_by(place) %>%
  summarize(ui_total = mean(ui_total, na.rm = TRUE))%>%
  ungroup()%>%
  select(place, ui_total, lf_total) %>%
  distinct(place, .keep_all = TRUE)

#divide total unemployed by working pop
fortable_unemp <- left_join(studytowns_UI, studytowns_LF, by = "place") %>%
  mutate(avgunemp_9 = (ui_total/lf_total)*100) %>%
  select(place, ui_total, lf_total, avgunemp_9) %>%
  distinct(place, .keep_all = TRUE)


```


#ES-202
##Import Data
```{r}
ES202_path <-"J:/EPPR/SEIGMA/Secondary Data/Census Data/Kassie/censusdata/ES202_All Muni_2017-2021/"
ES202_workbooks <- list.files(ES202_path)
workbook_names <-stringr::str_remove(ES202_workbooks, ".xlsx")

process1 <- lapply(paste(ES202_path, ES202_workbooks, sep = ""), read_excel)
names(process1) <- workbook_names

ES202 <- bind_rows(process1)

```

##Clean Data

```{r}
ES202 <- left_join(ES202, study_comm, by = c("areaname" = "ES202_match"))%>%
  filter(naicstitle == "Total, All Industries" | naicstitle == "Manufacturing")%>%
  filter(lvl != "2")

```


###10. Percent of WF Employed in Manufacturing
For MA Towns
```{r}
MAtowns_manuf <- ES202 %>%
  select(periodyear, place, naicstitle, avgemp) %>%
  pivot_wider(
    names_from = naicstitle,
    values_from = avgemp,
    names_prefix = "emp_",
    values_fill = 0
  ) %>%
  rename(total = "emp_Total, All Industries",
         manuf = "emp_Manufacturing",
         year = "periodyear") %>%
  group_by(place) %>%
  summarize(manuf_avg = mean(manuf),
            total_avg = mean(total))%>%
  group_by(place)%>%
  summarize(pctmanuf_10 = (manuf_avg / total_avg) * 100) %>%
  select(place, pctmanuf_10)


missing_data <- data.frame(
  place = c("Middlefield, MA", "Monroe, MA", "Mount Washington, MA"),
  pctmanuf_10 = rep(NA, 3)
 )

MAtowns_manuf <- rbind(MAtowns_manuf, missing_data) %>%
  distinct(place, .keep_all = TRUE)

```

###11. Average Annual Job Growth Rate
For MA Towns
```{r}
MAtowns_jobgrowth <- ES202 %>%
  select(periodyear, place, naicstitle, avgemp) %>%
  pivot_wider(
    names_from = naicstitle,
    values_from = avgemp,
    names_prefix = "emp_",
    values_fill = 0
  ) %>%
  rename(total = "emp_Total, All Industries",
         year = "periodyear")%>%
  select(-emp_Manufacturing)%>%
  arrange(place, year) %>%
  group_by(place) %>%
  mutate(percent_change = ((total - lag(total)) / abs(lag(total))) * 100) %>% #(new value - old value)/|old value|
  ungroup()

MAtowns_jobgrowth <- MAtowns_jobgrowth %>%
  filter(year != 2017) %>%
  select(place, percent_change) %>%
  group_by(place)%>%
  summarize(avgjobgrowth_11 = mean(percent_change))%>%
  ungroup()

missing_data2 <- data.frame(
  place = c("Middlefield, MA", "Monroe, MA", "Mount Washington, MA"),
  avgjobgrowth_11 = rep(NA, 3)
 )

MAtowns_jobgrowth <- rbind(MAtowns_jobgrowth, missing_data2)%>%
  distinct(place, .keep_all = TRUE)

```



#LODES
##Import Data
```{r}
LODES_path <- "OnTheMap_StudyComm_2017-2021/AllTowns_OntheMapWORK_2017_2021.xlsx"
LODES <- read_excel(LODES_path)

```

##Clean Data

###10.2. Percent of WF Employed in Manufacturing
For other study comm
```{r}
studytowns_manuf <- LODES %>%
  pivot_wider(
    names_from = Measure,
    values_from = Count,
    names_prefix = "emp_",
    values_fill = 0
  ) %>%
  rename(totalAll = "emp_TotalJobs",
         manuf = "emp_ManufacturingJobs",
         place = Place)%>%
  group_by(place)%>%
  summarize(manuf_sum = sum(manuf),
            totalAll_sum = sum(totalAll))%>%
  group_by(place)%>%
  mutate(pctmanuf_10 = (manuf_sum/totalAll_sum)*100) %>%
  select(place, pctmanuf_10)%>%
  ungroup()

studytowns_manuf <- left_join(studytowns_manuf, study_comm, by = "place")%>%
  distinct(place, .keep_all = TRUE)

```


```{r}
#combine with ES202 data for full dataset
fortable_manuf <- full_join(studytowns_manuf, MAtowns_manuf, by = "place")%>%
    mutate(pctmanuf_10 = coalesce(pctmanuf_10.y, pctmanuf_10.x),
      # Add more variables as needed
      .keep = "unused"
    )%>%
  select(place, comm_type, state, GEOID, pctmanuf_10)%>%
  distinct(place, .keep_all = TRUE)
```

###11.2. Average Annual Job Growth Rate
For other study comm
```{r}
studytowns_jobgrowth <- LODES %>%
  pivot_wider(
    names_from = Measure,
    values_from = Count,
    names_prefix = "emp_",
    values_fill = 0
  ) %>%
  rename(totalAll = "emp_TotalJobs",
         manuf = "emp_ManufacturingJobs",
         place = Place,
         year = Year)%>%
  select(-manuf) %>%
  arrange(place, year) %>%
    group_by(place) %>%
    mutate(percent_change = ((totalAll - lag(totalAll)) / abs(lag(totalAll))) * 100) %>% #(new value - old value)/|old value|
    ungroup() 

studytowns_jobgrowth <- studytowns_jobgrowth %>%
  filter(year != 2017) %>%
  select(place, percent_change) %>%
  group_by(place)%>%
  summarize(avgjobgrowth_11 = mean(percent_change))%>%
  ungroup()

studytowns_jobgrowth <- left_join(studytowns_jobgrowth, study_comm, by = "place")%>%
  distinct(place, .keep_all = TRUE)
```

```{r}
#combine with ES202 data for full dataset
fortable_jobgrowth <- full_join(studytowns_jobgrowth, MAtowns_jobgrowth, by = "place") %>%
    mutate(avgjobgrowth_11 = coalesce(avgjobgrowth_11.y, avgjobgrowth_11.x),
      # Add more variables as needed
      .keep = "unused"
    )%>%
  select(place, comm_type, state, GEOID, avgjobgrowth_11)%>%
  distinct(place, .keep_all = TRUE)
```


#Merge

Start by creating a df that contains all variables for all communities. (done)
Then create a subset for each casino's host and match communities. (done)
Then create an avg for the match comm for each casino. (done)
Then create an average of each variable for all MA communities. (done) 

```{r}
fortable_dataframes <- ls(pattern = "^fortable_") #Identify data frames with the "fortable_" convention
fortable_list <- lapply(fortable_dataframes, get) #Create a list of data frames using lapply

#Print the structure and column names of each data frame in fortable_list
lapply(fortable_list, function(df) {
  cat("Data frame structure:\n")
  print(class(df))
  cat("Column names:\n")
  print(names(df))
  cat("\n")
})

left_join_by_place <- function(df1, df2) {
  left_join(df1, df2, by = "place")
}

fortable_merged <- reduce(fortable_list, left_join_by_place) %>%
  select(place, comm_type.y, state.y, GEOID.y, pop_1.y, pctbs_2, medianHHinc_3, pctblack_4, pcthispanic_5, pctpoverty_6, lfpr_8.2, avgunemp_9, pctmanuf_10, avgjobgrowth_11) %>%
  rename(comm_type = comm_type.y,
         state = state.y,
         GEOID = GEOID.y,
         pop_1 = pop_1.y) %>%
  mutate(
    comm_type = case_when(
      place == "Encore, MA" ~ "EBH_host",
      place == "Springfield, MA" ~ "MGM_host",
      place == "Plainville, MA" ~ "PPC_host",
      # Add more conditions as needed
      TRUE ~ comm_type  #Keep the original value if no condition is met
    )
  )

```


```{r}
#create a subset for each casino's host and match communities
MGM <- fortable_merged %>%
  filter(str_detect(comm_type, "^MGM")) %>%
  mutate(pcthostpop = pop_1 / first(pop_1[comm_type == "MGM_host"]),
         pcthostmedHHinc = medianHHinc_3 / first(medianHHinc_3[comm_type == "MGM_host"]))

fortable_MGM <-  MGM%>%
  filter(comm_type != "MGM_alt",
         comm_type != "MGM_old") %>%
  bind_rows(
    fortable_merged %>%
      filter(comm_type == "MGM") %>%
      summarise(across(where(is.numeric), ~mean(., na.rm = TRUE))) %>%
      mutate(place = "Average Match Community", comm_type = "MGM")
  )


PPC <- fortable_merged %>%
  filter(str_detect(comm_type, "^PPC")) %>%
  mutate(pcthostpop = pop_1 / first(pop_1[comm_type == "PPC_host"]),
         pcthostmedHHinc = medianHHinc_3 / first(medianHHinc_3[comm_type == "PPC_host"]))

fortable_PPC <-  PPC%>%
  filter(comm_type != "PPC_alt",
         comm_type != "PPC_old") %>%
  bind_rows(
    fortable_merged %>%
      filter(comm_type == "PPC") %>%
      summarise(across(where(is.numeric), ~mean(., na.rm = TRUE))) %>%
      mutate(place = "Average Match Community", comm_type = "PPC")
  )


EBH <- fortable_merged %>%
  filter(str_detect(comm_type, "^EBH")) %>%
  mutate(pcthostpop = pop_1 / first(pop_1[comm_type == "EBH_host"]),
         pcthostmedHHinc = medianHHinc_3 / first(medianHHinc_3[comm_type == "EBH_host"]))

fortable_EBH <-  EBH%>%
  filter(comm_type != "EBH_alt",
         comm_type != "EBH_old") %>%
  bind_rows(
    fortable_merged %>%
      filter(comm_type == "EBH") %>%
      summarise(across(where(is.numeric), ~mean(., na.rm = TRUE))) %>%
      mutate(place = "Average Match Community", comm_type = "EBH")
  )

fortable_MAavg <- fortable_merged %>%
  filter(comm_type %in% c("MA_avg", "EBH_host", "MGM_host", "PPC_host")) %>%
  summarise(across(where(is.numeric), ~mean(., na.rm = TRUE))) %>%
  mutate(place = "Average Massachusetts Community", comm_type = "Average")

```

```{r}

fortable <- bind_rows(fortable_EBH, fortable_MGM, fortable_PPC, fortable_MAavg)

```

# Export

```{r}

write.xlsx(fortable, "fortable_20240422.xlsx", rowNames = FALSE)


```

```{r}
#create a subset for each casino's host and match communities

fortable_MGMalt <-  MGM %>%
  filter(comm_type != "MGM",
         comm_type != "MGM_host") 

fortable_PPCalt <-  PPC%>%
  filter(comm_type != "PPC",
         comm_type != "PPC_host") 

fortable_EBHalt <-  EBH%>%
  filter(comm_type != "EBH",
         comm_type != "EBH_host") 

fortable_alt <- bind_rows(fortable_EBHalt, fortable_MGMalt, fortable_PPCalt)


```

```{r}
write.xlsx(fortable_alt, "fortable_alt_20240422.xlsx", rowNames = FALSE)
```

